---
title: "Getting Started with oncoAgent"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with oncoAgent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Overview

oncoAgent is a semi-autonomous research system for oncology target discovery
that integrates multi-omics data with agentic AI workflows.

## Quick Start

### 1. Create Pipeline Configuration

```{r}
library(oncoAgent)

config <- create_pipeline_config(
  project = "TCGA-BRCA",
  cancer_type = "breast cancer",
  de_methods = c("deseq2", "edger", "limma"),
  padj_threshold = 0.05,
  output_dir = "results"
)
```

### 2. Run with Pre-loaded Data

```{r}
# Using synthetic/demo data
counts <- matrix(rnbinom(100 * 20, mu = 100, size = 10),
                 nrow = 100, ncol = 20)
rownames(counts) <- paste0("Gene_", 1:100)
colnames(counts) <- paste0("Sample_", 1:20)

col_data <- data.frame(
  condition = factor(rep(c("tumor", "normal"), each = 10)),
  row.names = colnames(counts)
)

clinical <- data.frame(
  sample_id = colnames(counts),
  time = rexp(20, 0.01),
  status = rbinom(20, 1, 0.5)
)

result <- run_oncopipeline(
  config,
  existing_data = list(
    expression = counts,
    col_data = col_data,
    clinical = clinical
  )
)
```

### 3. Run with Real TCGA Data

```{r}
config <- create_pipeline_config(
  project = "TCGA-BRCA",
  cancer_type = "breast cancer"
)

result <- run_oncopipeline(config)
```

## Individual Modules

### Data Ingestion

```{r}
# Fetch TCGA data
tcga <- fetch_tcga_data("TCGA-BRCA")

# Fetch DepMap CRISPR data
crispr <- fetch_depmap_crispr()

# Fetch drug sensitivity
gdsc <- fetch_gdsc_drug_sensitivity()
```

### Differential Expression

```{r}
# Single method
de_deseq2 <- run_deseq2(counts, col_data)

# Consensus across methods
de_consensus <- consensus_de(counts, col_data,
                             methods = c("deseq2", "edger", "limma"))
```

### Survival Analysis

```{r}
hr_table <- calculate_hazard_ratios(
  counts,
  time = clinical$time,
  status = clinical$status
)
```

### Network Analysis

```{r}
genes <- rownames(counts)[1:50]
ppi <- fetch_string_ppi(genes)
network <- build_ppi_network(ppi)
centrality <- calculate_network_centrality(network)
modules <- detect_network_modules(network)
```

### Target Scoring

```{r}
scores <- composite_target_score(
  de_results = de_consensus,
  survival_hr = hr_table,
  centrality = centrality
)

ranked <- rank_targets(scores, top_n = 20)
```

## Agentic Workflow

The five agents work together:

```{r}
# Data Discovery Agent
dd_agent <- DataDiscoveryAgent$new(project = "TCGA-BRCA")
dd_agent$assess(existing_data)

# Analysis Agent
analysis_agent <- AnalysisAgent$new()
results <- analysis_agent$run(data)

# Gap Detection Agent
gap_agent <- GapDetectionAgent$new()
gaps <- gap_agent$detect(results, data)

# Hypothesis Agent
hyp_agent <- HypothesisAgent$new()
hypotheses <- hyp_agent$generate(results$scoring)

# Evaluation Agent
eval_agent <- EvaluationAgent$new()
metrics <- eval_agent$evaluate(ranked, gold_standard)
```

## Evaluation

```{r}
gs <- build_reference_panel(sources = c("cosmic", "cgc"))

metrics <- benchmark_vs_goldstandard(ranked, gs)
auroc <- calculate_auroc(ranked, gs$genes)
report <- generate_evaluation_report(ranked, gs, "eval_report.md")
```
